A thesis draft

### 2018-11-27 res/sections/08-Background.tex
Add final part about arch impl

\section{Virtualization}
It is remarkable how cloud computing has surged in the recent years. This is due
the numerous features it carries, such as a great way to isolate processes and
mitigate possible attacks from malicious users. On top of that, it allows to run
virtual machines with different configurations, without bringing any changes to
the host machine. Indeed, the virtualizations concepts refer to the property of
a machine to be virtualized. To achieve this, the components that forms a common
computer need to be virtualized too. Thus, these assets are recreated in this
virtualized manner, e.g. storage devices and computer network
resources~\cite{liu2014research}. At the time writing, there are two common
virtualization options: the virtualization through Hypervisor (defined also as
hypervisor-based) and the container virtualization.

@@

\section{Service Function Chaining (SFC)}
Another important component we have to talk about is the SFC, defined as a
concept that it collocates logically above the NFVs. As said in~\cite{rfc7665}:
\begin{displayquote}
The definition and instantiation of an ordered set of service functions and
subsequent ``steering'' of traffic through them is termed Service Function
Chaining (SFC).
\end{displayquote}

>>>>

\section{Virtualization}
It is remarkable how cloud computing has surged in the recent years. This is due
the numerous features it carries, such as a great way to isolate processes and
mitigate possible attacks from malicious users. On top of that, it allows to run
virtual machines with different configurations, without bringing any changes to
the host machine. Indeed, the virtualization concepts refer to the property of
a machine to be virtualized. To achieve this, the components that forms a common
computer need to be virtualized too. Thus, these assets are recreated in this
virtualized manner, e.g. storage devices and computer network
resources~\cite{liu2014research}. At the time writing, there are two common
virtualization options: the virtualization through Hypervisor (defined also as
hypervisor-based) and the container virtualization.

@@

\section{Service Function Chaining (SFC)}
\label{chap:background:sec:sfc}
Another important component we have to talk about is the SFC, defined as a
concept that it collocates logically above the NFVs. As said in~\cite{rfc7665}:
\begin{displayquote}
The definition and instantiation of an ordered set of service functions and
subsequent ``steering'' of traffic through them is termed Service Function
Chaining (SFC).
\end{displayquote}

### 2018-11-27 res/sections/15-ArchitecturalImplementation.tex
Add final part about arch impl

The impossibility to develop a custom VIM driver and VNFM manager lead us to
search for an alternative solution. Then, we started to work directly on the
VIM. The base approach was to create an API Server which in future an Openbaton
plugin could contact and simply deploy TOSCA definitions, interacting with a set
of RESTful APIs.

% TODO talk about the problems of mapping the TOSCA definition from Openbaton
% with the Kubernetes elements (which we actually solved), and talk about how we
% weren't able to develop a VNFM plugin for Openbaton, and describe how Harbor
% born from it.

>>>>

The impossibility to develop a custom VIM driver and VNFM manager lead us to
search for an alternative solution. Then, we started to work directly on the
VIM. The base approach was to create an API Server which in future an Openbaton
plugin could contact and simply deploy TOSCA definitions, interacting with a set
of RESTful APIs. We called this simple backend \emph{Harbor}.

\subsection{Harbor}
\label{chap:archimpl:sec:secondattempt:sub:harbor}
Harbor grew considerably during the project development, as it revealed to be a
key component of our architecture. Harbor does not have a web interface, but as
already written it has a RESTful API that allows users to interact with it using
CRUD operations. It offers functionalities similar to Openbaton: possibility to
save and retrieve VNFs and SFCs definitions. When an SFC is saved, a consistency
check is performed to assure that all the VNFs have already been defined in
Harbor. On an SFC deployment, with a mechanism similar to the Openbaton one, all
the associated VNFs are set up too, making the whole operation of launching a
SFC atomic from the user point-of-view. Since this architecture component grew
until it became a Kubernetes MANO, we added a feature that, in our opinion,
Openbaton lacks: smart resource utilization. Harbor, in fact, during an SFC
deletion, it checks if any related VNF remains orphan (which means without any
SFC using it). In that case the VNF will enter in a state of pruning: after five
minutes, if the VNF has not be re-utilized by any other SFC it will be deleted.
This mechanism assure faster deployments in environments where NSs are
continuously taken up and down.

\begin{figure}[t]
 \centering
 \includegraphics[scale=0.5]{Harbor_architecture}
 \caption[Harbor architectural schema]{Harbor architectural schema}
 \label{chap:archimpl:sec:secondattempt:img:harborarchitecture}
\end{figure}

\subsection{The need for additional components}

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.5]{Attempt2v1Harbor}
  \caption[Second attempt revision one schema]{Second attempt revision one
    schema. It is possible to see how Harbor completely substituted Openbaton.
    Additionally, Harbor now operates directly with the Kubernetes API,
    retrieving informations about the VNFs status from it. This causes the VNFs
    to be lighter because they don't have to include any reporting tools, like
    the EMS, in it.}
  \label{chap:archimple:sec:secondattempt:img:attempt2v1harbor}
\end{figure}

Terminated the Harbor development, we ended up in a situation similar the one
depicted in Figure~\ref{chap:archimple:sec:secondattempt:img:attempt2v1harbor}.
It is possible to see how the Openbaton framework is not necessary anymore:
Harbor, in fact, ended up substituting it. We then started to think about the
challenges of integrating the SFC structure inside a Kubernetes cluster. We
identified the following challenges:
\begin{itemize}
\item whole system scalability: we wanted that every component had the
  possibility to scale during intense workloads, without the necessity of human
  intervention.
\item ingress and egress points: there should be at least two points in the
  cluster were the packect enters and exits;
\item dynamic package routing: as formulated by one of our requirements, packet
  routing had to be dynamic. We immediatly identified a possibility to create
  this functionality with the already-built features of Kubernetes;
\item a framework to automatically forward the data elaborated by the VNF: we
  identified the possibility to build a framework that automatically handled the
  packet receiving from another VNF, it properly called the function that had to
  elaborate the packet payload and then it forwarded the result to the next VNF;
\item connection status: having ingress and egress points implicates the
  necessity to save some persistent data in a distributed manner.
\end{itemize}

\subsubsection{Scalability considerations}
When adopting Kubernetes, we realized the possibility to make every component
easily scalable. Indeed, all the following components can be described in
Kubernetes as a \verb!Service!. For a \verb!Service! a \verb!ReplicaSet! and a
consequent \verb!Deployment! can be associated, giving the possibility for it to
scale under heavy workloads. The introduction of scalability meant that was not
possible to refer to the components based on their IP, but we had to refer to
their service name. As it will be explained later, this introduced challenges
especially for the ingress and egress points of our network, that have to keep a
connection with the sender and the receiver. On top of that, the scalability and
the required DNS lookup meant a possible bottleneck. With a great number of
components and VNFs constantly quering the centralized Kubernetes DNS lookup
system a performance degradation could have easily happened. We decided to
distribute the DNS service too, integrating the CoreDNS Kubernetes default
service with \emph{DNSMasq}, a DNS server able to perform query caching for
faster name resolutions. We created a DNSMasq \verb!DaemonSets!, which in turn
deployed one Pod with DNSMasq running inside for every minion in the cluster. At
the time of the VNF launch, we set that the DNS resolution should pass first for
the local \verb!DNSMasq! DNS, and then for the Kubernetes CoreDNS service. In
this way, we were able to distribute the DNS query workload from the master,
using a local caching system to speed up the name resolution between services.

\subsubsection{Ingress and Egress: Ironhide}
As just said, the whole framework needs to have and entry and an exit point.
These, defined as \emph{Ingress} and \emph{Egress} points, are performed by
\emph{Ironhide}. Ironhide acts like a hides the whole packet processing to the
sender $S$ and the receiver $R$. When a client $S$ performs a request to $R$,
the packet actually goes through Ironhide, who has an internal classifier that
it is able to recognize the type of connection and to choose the right NS the
packet has to go through. This process is fundamental since it keeps the
connection between $S$ and $R$ up, for both TCP and UDP based communications.
Since the connection need to be end-to-end, the state of the connections are
saved in the \emph{Roulette Controller} component.

\subsubsection{Dyamic package routing: the Roulette component}
\label{chap:prjan:sec:tech:sub:SFC:sub:roulette}
Dynamic end-to-end communications between VNFs would not be possible without a
third component acting as a route controller. Roulette fulfills this role,
providing a back-end where the single VNF can ask the next packet hop. Indeed,
routes must be submitted by another component, that we identified as Harbor.
After a NS deployment, Roulette gets instructed with a new route describing a
NS. Every NS has a proper ID (called \verb!sfid!), and thus when the classifier
checks a packet typology, it labels that packet with the specific NS ID (for a
detailed description about the SFC header see
Section~\ref{chap:background:sec:sfc}). Finally, thanks to this label, every VNF
knows the packet's chain and it will correctly forward the it to the next VNF.

\subsubsection{Keeping the connection status}
The scalability challenge introduces new issues in the system. The most
important is regarding keeping up stateful connections, since the ingress and
egress points could be more than one container. With this considerations in
mind, we started thinking at how and where keeping information about the ingress
container handling the connection with the sender $S$ and about the egress
container talking with the receiver $R$. Since the Roulette component has a
complete end-to-end topology view of the system, we decided to store these
additional informations (we called it \emph{endpoints}) in it.

\subsubsection{A VNF framework: Astaire}
As is possible to denote, a VNF doesn't only process the incoming packets, but
it has to manage the communications and the related forward operations. To
possibly ease part of the transmission problems and to make it easier to create
a VNF without requiring the whole knowledge about SFC, we decided to build a
framework, Astaire, that could kept the SFC related details decoupled from the
core of the VNF, in other words, the entity that actually elaborates the packet.
Built using an event-based pattern, it presents hooks that allow other software
written in different languages from the Astaire one (e.g. Java), to be called
dynamically at runtime in order to process the packet. The result of this
elaboration gets returned to Astaire, that performs the beforehand cited
operations of VNFs communications and forwarding. In the SFC specification,
Astaire is defined as a Service Function Forwarder (SFF).

\subsection{Final architecture structure}
At the end, we were able to create a scalable architecture that followed the
ETSI MANO standards, with Kubernetes as VIM and Harbor has MANO. We faced
additional challenges regarding the system scalability, that we solved using the
Roulette controller. We eased the creation of VNFs with Astaire, and we created
proper ingress and egress points with Ironhide, that supported scalability
issues and performed a protocol-based packet classification. Is possible to see
a final architectural schema in
Figure~\ref{chap:archimpl:sec:secondattempt:img:attempt2v1total}.

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.48]{Attempt2v1Total}
  \caption[Final components organization]{Final components organization. In the
    image is possible to denote how Egress, Ingress, the Roulette Controller and
    the VNFs instances are optional or could not be present in the minion, while
    DNSMasq is always deployed.}
  \label{chap:archimpl:sec:secondattempt:img:attempt2v1total}
\end{figure}
